Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-04-27 ms.tex

\begin{figure}
\includegraphics[width=\columnwidth]{mdot_sfr.pdf}
\caption{\label{fig:bh_growth} {\it Top panel:} Gas inflow rate
  (excluding small-scale outflows) $\dot{M}$ in Eddington units as a
  function of black hole mass $M_{\bullet}$, shown separately for core
  ($\Gamma=0.1$, dot-dashed) and cusp galaxies ($\Gamma=0.8$, solid).
  Black lines show the accretion rate calculated from equation
  (\ref{eq:eddr_analytic}) using our estimate of the gas heating rate
  provided at $z = 0$ by the average star formation histories for each
  galaxy mass (Fig.~\ref{fig:vwSources}).  Teal lines show the maximum
  accretion set by Ia supernovae blow-out (\ref{eq:eddr_Ia}).  Red
  lines show the accretion rate obtained if Compton heating alone was
  acting (eq.~[\ref{eq:MdotC}]).  Yellow lines show the maximum
  accretion rate for a thermally stable flow near the stagnation
  radius (eq.~[\ref{eq:Mdotmax}]).  We also show Eddington ratios
  inferred for SgrA* from \citet{Quataert:2004a} (gray circles), for
  NGC3115 from \citet{ShcherbakovWong+:2014a} (gray diamonds), and for
  the Swift J1644+57 TDE from \citet{BergerZauderer+:2012a} (gray
  triangles). Note that the actual black hole accretion rate could one
  to several orders of magnitude smaller due to the presence of small
  scale outflows. Thus the plot is really showing
  the gas inflow rate on large scales (e.g. near the stagnation
  radius) {\it Bottom panel:} Growth times $\Mbh/\Mdot$ in units of
  the Hubble time $\th$ for each of the accretion rates shown in the
  top panel.}
\end{figure}

@@

For low mass black holes, blow out from SN Ia radius caps the SMBH
accretion rate at a value $\dot{M}_{\rm Ia} \sim
10^{-6}-10^{-5}M_{\odot}$ yr$^{-1}$ (green lines;
eq.~[\ref{eq:eddr_Ia}]), which is however not low enough to prevent
otherwise thermally-unstable flow at smaller radii (eq.~[\ref{eq:heatingratioIa}]).  On the other hand, for higher
$M_{\bullet}$ the stagnation resides close to the Ia radius so Ia
heating contributes to the steady gas heating rate, even if the energy
released by a single Ia is insufficient to unbind gas from the stellar
bulge (eq.~[\ref{eq:blowout}]).

@@

Shown also for comparison in Figure \ref{fig:bh_growth} (top panel) is
the accretion rate onto SgrA* calculated by \citet{Quataert:2004a}.
Also show is the range of $\dot{M}$ for the low-luminosity AGN NGC3115
derived through detailed modeling by \citealt{ShcherbakovWong+:2014a},
who find a range of accretion rates depending on the assumed model for
thermal conductivity (although we expect this range would be smaller
had they included saturation of the conductive flux).  The SgrA* rate
exceeds our estimate of $\dot{M}$ due to the average star formation
history for the same $M_{\bullet}$ by almost two orders of magnitude.
This discrepancy can be understood by the fact that star formation in
the central parsec of the Milky Way cannot be described as
steady-state: feedback is instead dominated by the stellar winds from
the ring of young massive stars of mass $\sim 10^{4}M_{\odot}$ and
estimated age $\sim 10$ Myr (e.g., \citealt{Schodel+07}).  Such a star
formation history is better described by our `star burst' scenario,
for which the value of $\eta \sim 100$ at $\tau_{\star} = 10^{7}$ yr
(Fig.~\ref{fig:vwImp}) is a factor of $\sim 100$ times higher than the
value of $\eta \sim 1$ predicted in the average SFH case
(Fig.~\ref{fig:vwSources}, bottom panel).

@@

The left panel of Figure \ref{fig:miller} is calculated assuming a
constant low value for the radiative efficiency of $\epsilon_X =
10^{-4}$, typical of those estimated for low luminosity AGN (e.g.,
\citealt{Ho:2009a}), while the right panel luminosities are calculated
instead assuming an efficiency of $\epsilon_X= 10 \epsilon_{\rm
  bol}\dot{M}/\dot{M}_{\rm edd}$ that depends on the Eddington ratio
as predicted by some models for radiatively inefficient accretion
flows (\citealt{Narayan&Yi95}; \citealt{Narayan+98}).  Shown for
comparison are the X-ray measurements (green stars) and upper limits
(gray triangles) from the sample of early-type galaxies compiled by
\citet{Miller+15} (cf.~\citealt{Gallo+10}).  A green line shows the
best power-law fit to the X-ray luminosity from \citet{Miller+15},
given by $\langle L_X/L_{\rm edd}\rangle \sim \Mbh^\alpha$ with
$\alpha = -0.2$ (see also \citealt{Zhang+09, Pellegrini10, Gallo+10}).
Also shown are the maximum accretion rates, respectively, for
thermally stable accretion (eq.~[\ref{eq:Mdotmax}]; yellow), as set by
SN Ia blow-out (eq.~[\ref{eq:eddr_Ia}]; teal), and as allowed by
Compton heating feedback (eq.~[\ref{eq:MdotC}]; red).

\begin{figure*}
\includegraphics[width=\textwidth]{miller.pdf}
\caption{\label{fig:miller} Average nuclear X-ray luminosity, $\langle
  L_X \rangle = \epsilon_X \Mdot c^2$, as a function of SMBH mass.
  Green lines show our prediction in the case of heating due to the
  average SFH for galaxies corresponding to each black hole mass
  (Fig.~\ref{fig:bh_growth}), calculated using a power-law fit for the
  relationship $\Gamma(M_{\bullet})=0.3 \Mbheight^{-0.24}$ derived from the
  \citet{LauerFaber+:2007a} sample.  Shown for comparison are the
  measurements ({\it black stars}) and upper limits ({\it gray
    triangles}) for $L_X/L_{\rm edd}$ values, from the
  \citet{Miller+15} sample of early-type galaxies (a black line shows
  the best power-law fit $\langle L_X/L_{\rm edd}\rangle \propto
  M_{\bullet}^{\alpha}$, with $\alpha = -0.2$).  The left panel is
  calculated assuming a constant radiative efficiency
  $\epsilon_X=10^{-4} (\epsilon_{\rm bol}/0.1)$, while the right hand
  panel assumes $\epsilon_X= 0.01 (\epsilon_{\rm bol}/0.1)
  (\alpha/0.1) \epsilon_{rad} ( (\alpha/0.1) 0.1 \eddr) $, where
  $\epsilon_{\rm rad}$ is given by equation~\eqref{eq:efficiency} and
  $\eddr$ corresponds to the gas inflow rate on large scales.  Shown
  for comparison are the X-ray luminosities calculated for the maximum
  thermally-stable accretion rate ({\it dashed orange line};
  eq.~[\ref{eq:Mdotmax}]); the SN Ia-regulated accretion rate ({\it
    dashed teal line}; eq.~[\ref{eq:eddr_Ia}]), and the Compton
  heating-regulated accretion rate ({\it dashed pink line};
  eq.~[\ref{eq:MdotCa}]), again all calculated for the average SFH
  corresponding to each SMBH mass.}
\end{figure*}

>>>>

\begin{figure}
\includegraphics[width=\columnwidth]{mdot_sfr.pdf}
\caption{\label{fig:bh_growth} {\it Top panel:} Gas inflow rate $\dot{M}/\dot{M}_{\rm edd}$ as a
  function of black hole mass $M_{\bullet}$, shown separately for core
  ($\Gamma=0.1$, dot-dashed) and cusp galaxies ($\Gamma=0.8$, solid).
  Black lines show the accretion rate calculated from equation
  (\ref{eq:eddr_analytic}) using our estimate of the gas heating rate
  provided at $z = 0$ by the average star formation histories for each
  galaxy mass (Fig.~\ref{fig:vwSources}).  Teal lines show the maximum
  accretion set by Ia supernovae blow-out (\ref{eq:eddr_Ia}).  Red
  lines show the accretion rate obtained if Compton heating alone was
  acting (eq.~[\ref{eq:MdotC}]).  Yellow lines show the maximum
  accretion rate for a thermally stable flow near the stagnation
  radius (eq.~[\ref{eq:Mdotmax}]).  We also show Eddington ratios
  inferred for SgrA* from \citet{Quataert:2004a} (gray circles), for
  NGC3115 from \citet{ShcherbakovWong+:2014a} (gray diamonds), and for
  the Swift J1644+57 TDE from \citet{BergerZauderer+:2012a} (gray
  triangles). Note that the actual black hole accretion rate could one
  to several orders of magnitude smaller due to the presence of small
  scale outflows. Thus the plot is really showing
  the gas inflow rate on large scales (e.g. near the stagnation
  radius) {\it Bottom panel:} Growth times $\Mbh/\dot{M}_{\bullet}$ in units of
  the Hubble time $\th$ for each of the accretion rates shown in the
  top panel, where $\dot{M}_{\bullet} = 0.1\dot{M}$ to account for the fraction of the inflowing matter lost to outflows from the accretion disk.}
\end{figure}

@@

For low mass black holes, blow out from SN Ia radius caps the SMBH
accretion rate at a value $\dot{M}_{\rm Ia} \sim
10^{-6}-10^{-5}M_{\odot}$ yr$^{-1}$ (green lines;
eq.~[\ref{eq:eddr_Ia}]), which is however not low enough to prevent
otherwise thermally-unstable flow at smaller radii (eq.~[\ref{eq:heatingratioIa}]).  For higher
$M_{\bullet}$ the stagnation resides close to the Ia radius so Ia
heating contributes to the steady gas heating rate, even if the energy
released by a single Ia is insufficient to unbind gas from the stellar
bulge (eq.~[\ref{eq:blowout}]).  This explains why the black and yellow lines mean at high $M_{\bullet}$.

@@

Shown also for comparison in Figure \ref{fig:bh_growth} (top panel) is
the accretion rate onto SgrA* calculated by \citet{Quataert:2004a}.
Also show is the range of $\dot{M}$ for the low-luminosity AGN NGC3115
derived through detailed modeling by \citet{ShcherbakovWong+:2014a},
who find a range of accretion rates depending on the assumed model for
thermal conductivity\footnote{We expect this range would be have been smaller
had \citet{ShcherbakovWong+:2014a} included saturation of the conductive flux}.  The SgrA* rate
exceeds our estimate of $\dot{M}$ due to the average SFH for the same $M_{\bullet}$ by almost two orders of magnitude.
This discrepancy can in part be understood by the fact that star formation in
the central parsec of the Milky Way cannot be described as
steady-state: feedback is instead dominated by the stellar winds from
the ring of young massive stars of mass $\sim 10^{4}M_{\odot}$ and
estimated age $\sim 10$ Myr (e.g., \citealt{Schodel+07}).  Such a star
formation history is better described by our impulsive (starburst) scenario,
for which the value of $\eta \sim 100$ at $\tau_{\star} = 10^{7}$ yr
(Fig.~\ref{fig:vwImp}) is a factor of $\sim 100$ times higher than the
value of $\eta \sim 1$ predicted in the average SFH case
(Fig.~\ref{fig:vwSources}, bottom panel).

@@

The left panel of Figure \ref{fig:miller} is calculated assuming a
constant low value for the radiative efficiency of $\epsilon_X =
10^{-4}$, typical of those estimated for low luminosity AGN (e.g.,
\citealt{Ho:2009a}), while the right panel luminosities are calculated
instead assuming an efficiency of $\epsilon_X= \epsilon_{\rm
  bol}\epsilon_{\rm rad}$ that depends on the Eddington ratio
as predicted by \citet{Sharma+2007} (see eq.~[\ref{eq:efficiency}] and surrounding discussion).  Shown for comparison are the X-ray measurements (green stars) and upper limits
(gray triangles) from the sample of early-type galaxies compiled by
\citet{Miller+15} (cf.~\citealt{Gallo+10}).  A green line shows the
best power-law fit to the X-ray luminosity from \citet{Miller+15},
given by $\langle L_X/L_{\rm edd}\rangle \sim \Mbh^\alpha$ with
$\alpha = -0.2$ (see also \citealt{Zhang+09, Pellegrini10, Gallo+10}).
Also shown are the maximum accretion rates, respectively, for
thermally stable accretion (eq.~[\ref{eq:Mdotmax}]; yellow), as set by
SN Ia blow-out (eq.~[\ref{eq:eddr_Ia}]; teal), and as allowed by
Compton heating feedback (eq.~[\ref{eq:MdotC}]; red).

\begin{figure*}
\includegraphics[width=\textwidth]{miller.pdf}
\caption{\label{fig:miller} Average nuclear X-ray luminosity, $\langle
  L_X \rangle = \epsilon_X \Mdot c^2$, as a function of SMBH mass.
  Green lines show our prediction in the case of heating due to the
  average SFH for galaxies corresponding to each black hole mass
  (Fig.~\ref{fig:bh_growth}), calculated using a power-law fit for the
  relationship $\Gamma(M_{\bullet})=0.3 \Mbheight^{-0.24}$ derived from the
  \citet{LauerFaber+:2007a} sample.  Shown for comparison are the
  measurements ({\it black stars}) and upper limits ({\it gray
    triangles}) for $L_X/L_{\rm edd}$ values, from the
  \citet{Miller+15} sample of early-type galaxies (a black line shows
  the best power-law fit $\langle L_X/L_{\rm edd}\rangle \propto
  M_{\bullet}^{\alpha}$, with $\alpha = -0.2$).  The left panel is
  calculated assuming a constant radiative efficiency
  $\epsilon_X=10^{-4} (\epsilon_{\rm bol}/0.1)$, while the right hand
  panel assumes $\epsilon_X= \alpha \epsilon_{\rm bol}\epsilon_{\rm rad}$, where $\epsilon_{\rm rad}$ is the radiative efficiency calculated by \citet{Sharma+2007} (eq.~[\ref{eq:efficiency}]) and $\alpha = 0.1$ is the fraction of the mass inflowing on large scales that reaches the SMBH.  Shown
  for comparison are the X-ray luminosities calculated for the maximum
  thermally-stable accretion rate ({\it dashed orange line};
  eq.~[\ref{eq:Mdotmax}]); the SN Ia-regulated accretion rate ({\it
    dashed teal line}; eq.~[\ref{eq:eddr_Ia}]), and the Compton
  heating-regulated accretion rate ({\it dashed pink line};
  eq.~[\ref{eq:MdotCa}]), again all calculated for the average SFH
  corresponding to each SMBH mass.}
\end{figure*}
